---
documentclass: article
fontsize: 12pt
date: "`r Sys.Date()`"
book_filename: "thesis_raisa_carmen"
delete_merged_file: TRUE
output: 
  bookdown::pdf_document2: 
    fig_caption: yes
    toc: false
    latex_engine: xelatex
    keep_tex: yes
    includes:
      in_header: preamble.sty
      before_body: titlepage.sty
      after_body: back_cover.sty
bibliography: references.bib  
editor_options: 
  markdown: 
    wrap: 72
site: bookdown::bookdown_site
---

```{r setup, include=FALSE}
library(frictionless) # to load data from the data package
library(ggstats) # for better likert plots
library(here)
library(kableExtra)
library(likert)
library(nFactors)
library(patchwork)
library(psych)
library(rprojroot)
library(tidyverse)
library(lavaan)
library(effectsize) 
library(semPlot)
library(semTools)
source(find_root_file("source", "R", "fa_table.R",
                       criterion = has_file("baseball_thesis.Rproj")))

knitr::opts_chunk$set(echo = FALSE)

baseball_package <- read_package(
  find_root_file(
    "data", "processed", "datapackage_analysis",
    "datapackage.json", 
    criterion = has_file("baseball_thesis.Rproj")))
atop_key <- read_resource(package = baseball_package,
              resource_name = "wp5_atop_questions")
nc_key <- read_resource(package = baseball_package,
              resource_name = "wp5_nc_key")
atop_key_p <- read_resource(package = baseball_package,
              resource_name = "wp5_parents_outdoor_play_key")
reps_key <- read_resource(package = baseball_package,
              resource_name = "wp5_parents_risk_protection_key")
im_key <- read_resource(package = baseball_package,
              resource_name = "wp5_parents_independent_mobility_key")
children <- read_resource(package = baseball_package,
              resource_name = "wp2_participants_metadata") %>%
  left_join(read_resource(package = baseball_package,
                          resource_name = "wp2_school_data") %>%
              dplyr::select(-n_students))
```

# Preface {-}

<!-- Preface: The preface should summarise the general aim of the work. People, offices and companies should be thanked for their support.  -->

B@seball (Biodiversity at School Environments Benefits for ALL) is a [\text{\underline{BELSPO project}}](http://www.belspo.be/belspo//brain2-be/projects/BatSEBALL_N.pdf) project that aims to link the general physical and mental health of children to the amount of greenery at the school environment. The general aim of this thesis is to analyse the scale questions that were included in the survey data from the B@seball project. To do this, confirmatory factor analysis and a structural equation model are applied.

Personally, I would like to thank;

My promotor, Prof. Michel Meulders, for the guidance and reflections provided during the writing of this thesis.

All B@seball project members, for allowing me to work with the data. A special thanks goes to the B@seball project members closest to me, Hans, Els, and Michael.

My employer, INBO, and my dear colleagues who granted me the time to work on the thesis, even during working hours.

My partner, Stefan, for proof-reading the thesis and for supporting me not only during the writing of the thesis but throughout the whole Master's program.

My family and friends for endless trust, support, and the necessary breaks from work.



\newpage
# Summary {-}

<!-- Add popularising summary here. (max. 2 pages for each language): The summary should present the most important aims and conclusions of the dissertation in a brief text of maximum two pages, written both in Dutch and in English for a Dutch Master programme, and only in English for an English Master programme. -->


The data used in this work originates from the [B@seball](https://www.uantwerpen.be/nl/projecten/baseball/) (Biodiversity at School Environments Benefits for ALL) project which is a BELSPO project (BRAIN-be 2.0 call). Project partners are the University of Antwerp, Sciensano, Goodplanet, UC Louvain, MOS, and INBO. The main aim of the project is to investigate whether and how biodiversity in school environments can influence the physical and mental health of children.
For this purpose, over 500 children between 10 and 12 years old from 36 different schools in Belgium were recruited.

Among many other health and mental development parameters, several standardized scale questions were introduced in surveys. These scales are based on existing scales (in English) from the literature. The scales were translated to Dutch and French to accommodate the survey respondents. These translated scales are the main focus ofthis report.

Children were first asked about their Attitude Towards Outdoor Play (ATOP) [@beyer2015development]. ATOP has two subscales, *ATOP_benefits* and *ATOP_fears*, that are measured on a likert scale by seven and four statements respectively. *ATOP_benefits* is high if the child values the benefits that outdoor play may have and *ATOP_fears* is high if the child is fearful of some elements of outdoor play. Second, the children's *Nature Connectedness* (NC) is measured by six affective statements, based on @hunt2017monitor.

Parents' tendency to (1) let children engage with risk and (2) prevent injuries is measures with two latent factors, *REPS_risk_engagement* and *REPS_injury_prevention*. These latent factors are based on six and eight statements respectively from the Risk Engagement and Protection Survey (REPS) [@olsen2018risk].

The first contribution of this report is analyzing these three scales (ATOP, REPS, and NC) and checking whether they reliably measure the underlying constructs. 
Using multi-group confirmatory factor analyses, we conclude that (1) NC can reliably measure a child's connection to nature, (2) ATOP can measure ATOP_benefits quite well but the ATOP_fears subscale still needs some improvement, (3) the two REPS subscales can currently not reliably measure the REPS_risk_engagement and REPS_injury_prevention concepts.

A second important contribution of this report is a Structural Equation Model (SEM) where we relate the ATOP subscales and NC to socio-economic status, gender, and the naturalness or greenness or the school environment to each other. We conclude that (1) children with high ATOP_benefits tend to have high NC, (2) children with a higher socio-economic status tend to have higher ATOP_benefits, ATOP_fears, and NC, (3) boys tend to have lower ATOP_benefits and higher ATOP_fears, and (4) children in schools with more natural school surroundings tend to have higher ATOP_benefits and NC.


\newpage

# List of abbreviations {-}

```{r abbreviations, purl = FALSE}
data.frame(
  Abbreviation = c("ATOP",
                   "CFA",
                   "CFI",
                   "EFA",
                   "NC",
                   "REPS",
                   "RMSEA",
                   "SEM",
                   "SES",
                   "SRMR",
                   "TLI"),
  Explanation = c("Attitude towards outdoor play",
                  "Confirmatory factor analysis",
                  "Comparative fit index",
                  "Exploratory factor analysis",
                  "Nature connectedness",
                  "Risk engagement and protection survey",
                  "Root mean square error of approximation",
                  "Structural equation model",
                  "Socio-economic status",
                  "Standardized root mean square residual",
                  "Tucker-Lewis index"
                  )
  ) %>%
  kable(booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = "hold_position")
```

\newpage
```{=latex}
% Trigger ToC creation in LaTeX
\tableofcontents
```
