---
documentclass: article
fontsize: 12pt
date: "`r Sys.Date()`"
book_filename: "thesis_raisa_carmen"
delete_merged_file: TRUE
output: 
  bookdown::pdf_document2: 
    fig_caption: yes
    toc: false
    latex_engine: xelatex
    keep_tex: yes
    includes:
      in_header: preamble.sty
      before_body: titlepage.sty
      after_body: back_cover.sty
bibliography: references.bib  
editor_options: 
  markdown: 
    wrap: 72
site: bookdown::bookdown_site
---

```{r setup, include=FALSE}
library(frictionless) # to load data from the data package
library(ggstats) # for better likert plots
library(here)
library(kableExtra)
library(likert)
library(nFactors)
library(patchwork)
library(psych)
library(rprojroot)
library(tidyverse)
library(lavaan)
library(effectsize) #to easily interpret SEM fit
library(semTools)
source(find_root_file("source", "R", "fa_table.R",
                       criterion = has_file("baseball_thesis.Rproj")))
knitr::opts_chunk$set(echo = FALSE)

baseball_package <- read_package(
  find_root_file(
    "data", "processed", "datapackage_analysis",
    "datapackage.json", 
    criterion = has_file("baseball_thesis.Rproj")))
atop_key <- read_resource(package = baseball_package,
              resource_name = "wp5_atop_questions")
nc_key <- read_resource(package = baseball_package,
              resource_name = "wp5_nc_key")
atop_key_p <- read_resource(package = baseball_package,
              resource_name = "wp5_parents_outdoor_play_key")
reps_key <- read_resource(package = baseball_package,
              resource_name = "wp5_parents_risk_protection_key")
im_key <- read_resource(package = baseball_package,
              resource_name = "wp5_parents_independent_mobility_key")
children <- read_resource(package = baseball_package,
              resource_name = "wp2_participants_metadata") %>%
  left_join(read_resource(package = baseball_package,
                          resource_name = "wp2_school_data") %>%
              dplyr::select(-n_students))
```

# Preface {-}

Preface: The preface should summarise the general aim of the work. People, offices and companies should be thanked for their support.  


\newpage
# Summary {-}

Add popularising summary here. (max. 2 pages for each language): The summary should present the most important aims and conclusions of the dissertation in a brief text of maximum two pages, written both in Dutch and in English for a Dutch Master programme, and only in English for an English Master programme.

\newpage

# List of abbreviations {-}

```{r abbreviations}
data.frame(
  Abbreviation = c("ATOP",
                   "CFA",
                   "EFA",
                   "NC",
                   "REPS",
                   "SEM"),
  explanation = c("Attitude towards outdoor play",
                  "Confirmatory factor analysis",
                  "Exploratory factor analysis",
                  "Nature connectedness",
                  "Risk engagement and protection survey",
                  "Structural equation model"
                  )
  ) %>%
  kable(booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = "hold_position")
```

\newpage
```{=latex}
% Trigger ToC creation in LaTeX
\tableofcontents
```
