---
title: "Exploring ATOP in kids for French and Dutch questionnaires separately"
author: "Raïsa Carmen"
date: '`r Sys.Date()`'
output: 
  bookdown::html_document2: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(frictionless)
library(nFactors)
library(rprojroot)
library(psych)
source(find_root_file("source", "R", "fa_table.R",
                       criterion = has_file("baseball_thesis.Rproj")))
baseball_package <- read_package(
  find_root_file(
    "data", "processed", "datapackage_analysis",
    "datapackage.json", 
    criterion = has_file("baseball_thesis.Rproj")))
atop_key <- read_resource(package = baseball_package,
              resource_name = "wp5_atop_questions")
atop_data <- read_resource(package = baseball_package,
              resource_name = "wp5_atop_data")
children <- read_resource(package = baseball_package,
              resource_name = "wp2_participants_metadata") %>%
  left_join(read_resource(package = baseball_package,
                          resource_name = "wp2_school_data") %>%
              dplyr::select(-n_students))
data_cfa_atop <- atop_data %>%
  dplyr::select(child_id, question_short, response_values) %>%
  mutate(response_values = ifelse(response_values == -9,
                                  NA,
                                  response_values)) %>%
  group_by(child_id, question_short) %>%
  summarize(response_values = mean(response_values, na.rm = TRUE)) %>% #there are some children who filled it in more than once -> we take the average score for them
  ungroup() %>%
  left_join(children)
```

# Dutch version

```{r analysis-nl, message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
data_cfa_atop_nl <- data_cfa_atop %>% 
  filter(str_detect(community, "Dutch speaking")) %>%
  dplyr::select(child_id, question_short, response_values) %>%
  pivot_wider(id_cols = child_id, names_from = question_short,
              values_from = response_values) %>%
  dplyr::select(child_id, atop_key$question_short)

#first, we need to standardize our variables
data_cfa_atop_unscaled_nl <- data_cfa_atop_nl
data_cfa_atop_nl[,-1] <- scale(data_cfa_atop_nl[,-1], center = TRUE,
                               scale = FALSE)

data_cfa_atop_nl <- data_cfa_atop_nl[complete.cases(data_cfa_atop_nl),]
covmat <- cov(data_cfa_atop_nl[,-1])
cormat <- cor(data_cfa_atop_nl[,-1])

#exploratory factor analysis on correlation matrix with 3 factors and oblique rotation
efa_obl3_nl <- fa(cormat, covar = FALSE, 3, rotate = "oblimin", fm = "ml",
                  n.obs = nrow(data_cfa_atop_nl))
efa_obl2_nl <- fa(cormat, covar = FALSE, 2, rotate = "oblimin", fm = "ml",
                  n.obs = nrow(data_cfa_atop_nl))
#list matrix of structure loadings
print(efa_obl3_nl$Structure, cutoff = 0, digits = 3)
#list matrix of residual correlations
round(efa_obl3_nl$residual, 3)
#compute factor scores after oblique rotation
scores_efaobl3_nl <- factor.scores(data_cfa_atop_nl[,-1], efa_obl3_nl,
                                   method = "tenberge")
tables_nl_3 <- fa_table(efa_obl3_nl,
                        title = "Factor analysis results for Dutch-speaking
                        children (3 factors).",
                        varlabels = atop_key$question_text_english)
tables_nl_2 <- fa_table(efa_obl2_nl,
                        title = "Factor analysis results for Dutch-speaking
                        children (2 factors).",
                        varlabels = atop_key$question_text_english)
# apply PCA
pca <- prcomp(data_cfa_atop_nl[,-1])
scree_nl <- screeplot(pca, type = "lines")#3 factors is likely better
ev <- eigen(cormat) # get eigenvalues
ap <- parallel(subject = nrow(data_cfa_atop_nl),
               var = ncol(data_cfa_atop_nl) - 1,
  rep = 100, cent = .05)
nS <- nScree(x = ev$values, aparallel = ap$eigen$qevpea)
pdf(find_root_file(
    "data", "figures",
    "scree_atop_nl.pdf", 
    criterion = has_file("baseball_thesis.Rproj")), width = 6.5, height = 4)
plotnScree(nS)
dev.off()


alpha_ATOP_benefits_nl <- alpha(data_cfa_atop_unscaled_nl %>%
                               left_join(children %>%
                                           dplyr::select(child_id,
                                                         community)) %>%
                               dplyr::filter(str_detect(community, "Dutch")) %>%
                               dplyr::select(2:8))
alpha_ATOP_fears_nl <- alpha(data_cfa_atop_unscaled_nl %>%
                            left_join(children %>%
                                        dplyr::select(child_id,
                                                      community)) %>%
                            dplyr::filter(str_detect(community, "Dutch")) %>%
                            dplyr::select(9:12))
```

There are `r nrow(unique(data_cfa_atop[data_cfa_atop$community == "Dutch speaking", "child_id"]))` children in a Dutch-speaking school and `r nrow(unique(data_cfa_atop[data_cfa_atop$community == "Dutch speaking.French speaking", "child_id"]))` children in bilingual schools that filled in the ATOP scale questions.
Among them, `r nrow(data_cfa_atop_nl)` children filled in all of the questions an can therefor be analyzed further.

The scree plot indicates that two (according to optimal coordinates method) or three (according to the acceleration factor method) should be best (Figure \@ref(fig:atop-scree-nl)).

```{r atop-scree-nl, warning=FALSE, message=FALSE, fig.cap="Analysis of the number of component or factors to retain in an exploratory principal component or factor analysis.", echo=FALSE, include=TRUE}
knitr::include_graphics(find_root_file(
    "data", "figures",
    "scree_atop_nl.pdf", 
    criterion = has_file("baseball_thesis.Rproj")))
```

## Two factors
In an EFA with oblique rotation, the original two groups of scale items immediately emerge, as shown in the tables below.
There is little correlation between the two factors but they can only explain `r 100*round(efa_obl2_nl$Vaccounted[3,2],3)`% of the total variance.

```{r atop-2fact-nl-1, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
tables_nl_2$ind_table
```

```{r atop-2fact-nl-2, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
tables_nl_2$f_table
```

The raw Chronbach alpha for the ATOP_benefits scale is `r round(unlist(unname(alpha_ATOP_benefits_nl$total["raw_alpha"])), 2)`. The first column in Table \@ref(tab: atop-kids-alpha-nl) shows that this might be improved if the statement "I like to make up my own games when I am outdoors in nature." is removed from the scale. This statement also has a very low correlation with the scale. 

The raw Chronbach alpha for the ATOP_benefits scale is `r round(unlist(unname(alpha_ATOP_fears_nl$total["raw_alpha"])), 2)`. The first column in Table \@ref(tab: atop-kids-alpha-nl) shows that this alpha value would always go down if any of the four items is removed.


```{r atop-kids-alpha-nl, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
alpha_ATOP_benefits_nl$alpha.drop["raw_alpha"] %>%
  cbind(alpha_ATOP_benefits_nl$item.stats[c("r.drop", "mean", "sd")]) %>%
  rbind(alpha_ATOP_fears_nl$alpha.drop["raw_alpha"] %>%
          cbind(alpha_ATOP_fears_nl$item.stats[c("r.drop",
                                              "mean", "sd")])) %>%
  mutate(question_short = row.names(.)) %>%
  left_join(atop_key %>% dplyr::select(question_short, question_text_english),
            by = ) %>%
  dplyr::select(question_text_english, raw_alpha, r.drop, mean, sd) %>%
  kable(booktabs = TRUE,
        escape = FALSE,
        caption = "Reliability if an item is dropped and item statistics.",
        digits = 3,
        col.names = c("item", "raw $\\alpha$", "item-scale corr.", "mean", "std. dev.")) %>%
  kableExtra::group_rows(start_row = 1,
                         end_row = nrow(alpha_ATOP_benefits_nl$alpha.drop),
                         group_label = "ATOP_benefits") %>%
  kableExtra::group_rows(start_row = 1 + nrow(alpha_ATOP_benefits_nl$alpha.drop),
                         end_row = nrow(alpha_ATOP_benefits_nl$alpha.drop) +
                           nrow(alpha_ATOP_fears_nl$alpha.drop),
                         group_label = "ATOP_fears") %>%
  kableExtra::kable_styling(font_size = 11) %>%
  kableExtra::column_spec(1, width = "3in") %>%
  kableExtra::column_spec(3, width = "0.7in")
```

## Three factors
In an EFA with oblique rotation and three factors, the original ATOP_fears group still remains but the ATOP_benefits group breaks down into two groups: one bigger group on more tangible benefits (thinking clearly, better health, exploring and learning) and a smaller group on benefits for the mood (feeling calm and free).

There is quite some interfactor correlation between the two ATOP_benefits factors (`r round(efa_obl3_nl$Phi[1, 3],3)`).
All three factors together now explain `r 100*round(efa_obl3_nl$Vaccounted[3,3],3)`% of the total variance.
```{r atop-3fact-nl-1, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
tables_nl_3$ind_table
```

```{r atop3fact-nl-2, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
tables_nl_3$f_table
```


# French version

```{r analysis-fr, message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
data_cfa_atop_fr <- data_cfa_atop %>% 
  filter(str_detect(community, "French speaking")) %>%
  dplyr::select(child_id, question_short, response_values) %>%
  pivot_wider(id_cols = child_id, names_from = question_short,
              values_from = response_values) %>%
  dplyr::select(child_id, atop_key$question_short)

#first, we need to standardize our variables
data_cfa_atop_unscaled_fr <- data_cfa_atop_fr
data_cfa_atop_fr[,-1] <- scale(data_cfa_atop_fr[,-1], center = TRUE,
                               scale = FALSE)

data_cfa_atop_fr <- data_cfa_atop_fr[complete.cases(data_cfa_atop_fr),]
covmat <- cov(data_cfa_atop_fr[,-1])
cormat <- cor(data_cfa_atop_fr[,-1])

#exploratory factor analysis on correlation matrix with 3 factors and oblique rotation
efa_obl3_fr <- fa(cormat, covar = FALSE, 3, rotate = "oblimin", fm = "ml",
                  n.obs = nrow(data_cfa_atop_fr))
efa_obl2_fr <- fa(cormat, covar = FALSE, 2, rotate = "oblimin", fm = "ml",
                  n.obs = nrow(data_cfa_atop_fr))

#list matrix of structure loadings
print(efa_obl3_fr$Structure, cutoff = 0, digits = 3)
#list matrix of residual correlations
round(efa_obl3_fr$residual, 3)
#compute factor scores after oblique rotation
scores_efaobl3_fr <- factor.scores(data_cfa_atop_fr[,-1], efa_obl3_fr,
                                   method = "tenberge")
tables_fr_3 <- fa_table(efa_obl3_fr,
                        title = "Factor analysis results for French-speaking
                        children (3 factors).",
                        varlabels = atop_key$question_text_english)
tables_fr_2 <- fa_table(efa_obl2_fr,
                        title = "Factor analysis results for French-speaking
                        children (2 factors).",
                        varlabels = atop_key$question_text_english)
# apply PCA
pca <- prcomp(data_cfa_atop_fr[,-1])
scree_fr <- screeplot(pca, type = "lines")#3 factors is likely better
ev <- eigen(cormat) # get eigenvalues
ap <- parallel(subject = nrow(data_cfa_atop_fr), var = ncol(data_cfa_atop_fr)-1,
  rep = 100, cent = .05)
nS <- nScree(x = ev$values, aparallel = ap$eigen$qevpea)
pdf(find_root_file(
    "data", "figures",
    "scree_atop_fr.pdf", 
    criterion = has_file("baseball_thesis.Rproj")), width = 6.5, height = 4)
plotnScree(nS)
dev.off()

alpha_ATOP_benefits_fr <- alpha(data_cfa_atop_unscaled_fr %>%
                               left_join(children %>%
                                           dplyr::select(child_id,
                                                         community)) %>%
                               dplyr::filter(str_detect(community, "French")) %>%
                               dplyr::select(2:8))
alpha_ATOP_fears_fr <- alpha(data_cfa_atop_unscaled_fr %>%
                            left_join(children %>%
                                        dplyr::select(child_id,
                                                      community)) %>%
                            dplyr::filter(str_detect(community, "French")) %>%
                            dplyr::select(9:12))

```

There are `r nrow(unique(data_cfa_atop[data_cfa_atop$community == "French speaking", "child_id"]))` children in a French-speaking school and `r nrow(unique(data_cfa_atop[data_cfa_atop$community == "Dutch speaking.French speaking", "child_id"]))` children in bilingual schools that filled in the ATOP scale questions.
Among them, `r nrow(data_cfa_atop_fr)` children filled in all of the questions an can therefor be analyzed further.


The scree plot indicates that two (according to optimal coordinates method) or three (according to the acceleration factor method) should be best (Figure \@ref(fig:atop-scree-fr)).

```{r atop-scree-fr, warning=FALSE, message=FALSE, fig.cap="Analysis of the number of component or factors to retain in an exploratory principal component or factor analysis.", echo=FALSE, include=TRUE}
knitr::include_graphics(find_root_file(
    "data", "figures",
    "scree_atop_fr.pdf", 
    criterion = has_file("baseball_thesis.Rproj")))
```

## Two factors
In an EFA with oblique rotation, the original two groups of scale items immediately emerge, as shown in the tables below.
There is little correlation between the two factors but they can only explain `r 100*round(efa_obl2_fr$Vaccounted[3,2],3)`% of the total variance.

```{r atop_2fact_fr_1, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
tables_fr_2$ind_table
```

```{r atop-2fact-fr-2, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
tables_fr_2$f_table
```


Both raw Chronbach alphas are slightly lower for the French version of the questionnaire then for the Dutch version: `r round(unlist(unname(alpha_ATOP_benefits_fr$total["raw_alpha"])), 2)` and `r round(unlist(unname(alpha_ATOP_fears_fr$total["raw_alpha"])), 2)` for the ATOP_benefits and ATOP_fears scale respectively.
For the ATOP_benefits scale, there are no statements that would increase $\alpha$ when removed though the statement "I like to make up my own games when I am outdoors in nature." come close and has low correlation with the scale.


```{r atop-kids-alpha-fr, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
alpha_ATOP_benefits_fr$alpha.drop["raw_alpha"] %>%
  cbind(alpha_ATOP_benefits_fr$item.stats[c("r.drop", "mean", "sd")]) %>%
  rbind(alpha_ATOP_fears_fr$alpha.drop["raw_alpha"] %>%
          cbind(alpha_ATOP_fears_fr$item.stats[c("r.drop",
                                              "mean", "sd")])) %>%
  mutate(question_short = row.names(.)) %>%
  left_join(atop_key %>% dplyr::select(question_short, question_text_english),
            by = ) %>%
  dplyr::select(question_text_english, raw_alpha, r.drop, mean, sd) %>%
  kable(booktabs = TRUE,
        escape = FALSE,
        caption = "Reliability if an item is dropped and item statistics.",
        digits = 3,
        col.names = c("item", "raw $\\alpha$", "item-scale corr.", "mean", "std. dev.")) %>%
  kableExtra::group_rows(start_row = 1,
                         end_row = nrow(alpha_ATOP_benefits_fr$alpha.drop),
                         group_label = "ATOP_benefits") %>%
  kableExtra::group_rows(start_row = 1 + nrow(alpha_ATOP_benefits_fr$alpha.drop),
                         end_row = nrow(alpha_ATOP_benefits_fr$alpha.drop) +
                           nrow(alpha_ATOP_fears_fr$alpha.drop),
                         group_label = "ATOP_fears") %>%
  kableExtra::kable_styling(font_size = 11) %>%
  kableExtra::column_spec(1, width = "3in") %>%
  kableExtra::column_spec(3, width = "0.7in")
```

## Three factors

In an EFA with oblique rotation and three factors, the original ATOP_fears group still remains but the ATOP_benefits group breaks down into two groups. A first, bigger group includes all ATOP_benefits items except imagination and learning new things. The second group separates out learning new things. 
There is quite some interfactor correlation between the two ATOP_benefits factors (`r round(efa_obl3_fr$Phi[1, 3],3)`).
All three factors together now explain `r 100*round(efa_obl3_fr$Vaccounted[3,3],3)`% of the total variance.
```{r atop-3fact-fr-1, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
tables_fr_3$ind_table
```

```{r atop-3fact-fr-2, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
tables_fr_3$f_table
```


